Data Management Plan

### 2021-11-30 application/dataManagementPlan_long_joaco.tex
Multiple modificiations

Software modules will comprise source files in a number of programming languages: C\#, Python, MATLAB and R.  These will be accompanied by extensive documentation (discussed below).

NEED: Approximate lines of code??? (order of magnitude is fine)

NEED: testing files, makefiles?, other stuff,

@@

% code
The source code for the core Bonsai ecosystem and the majority of extension
packages are hosted at \url{github.com/bonsai-rx} (MIT license with no
copyright and licencing restrictions).

@@

\subsubsection*{Team responsibilities}

@@

    \item[Goncalo Lopes] Bonsai related software development,
        software releases, organization of Bonsai workshops
        and co-supervision of research software engineer (RSE).

    \item[Joaquin Rapela] machine learning (ML) related software development,
        interaction with ML functionality methods developers, validating of ML
        functionality with behavioral and neural recordings and co-supervision
        of RSE.

    \item[Maneesh Sahani] advise on ML.

    \item[Tom Mrsic-Flogel] advise on neuroscience topics and communication
        with the steering committee.

\subsection*{Resourcing and responsibility}

    \item[software development model] we will use the Kanban software
        development model, which we both Dr.~Lopes and Dr.~Rapela are already
        using in a collaborative project at the SWC/GCNU.

    \item[weekly meetings] between Goncalo Lopes, Joaquin Rapela and the RSE
        will be an essential component of the software development process.
        These will ensure that every member of the team is informed about both
        the Bonsai integration and the machine learning developments and will
        make the software development team more reliable upon the absence of
        any of its members.

    \item[code review and pair-programming] sessions will be held on a monthly basis among Goncalo Lopes, Joaquin Rapela and the recruited software engineer to deliver readable code sharing a standard programming style.

\subsubsection*{Major risks}

    \item[language integration] we proposed to use existing software libraries
        to enable the communication between Bonsai, written in C\#, and other
        machine learning algorithms, written in Python (Python.NET), R (R.Net)
        and Matlab (native Matlab .Net interface). We have not yet tested any
        of these existing libraries, and we may found that they don't function
        reliably enough for our purposes. If this were the case, we proposed to
        build our own interface protocol using the ZeroMQ messaging library.
        This risk will only delay the progress of the proposed project, but
        will not hinder the deployment of its proposed outputs.

    \item[disfunctional machine learning functionality] in Section 2.3 of the
        Case for Support document (Figure 2, green box) we proposed to
        integrate existing machine learning algorithms into the Bonsai
        ecsosystem. It could be the case that some of these algorithms do not
        work as reliably as needed for our resource. I such a case we will use
        another algorithm with similar functionality.  This risk should not be
        a major problem, since most of the machine learning functionality we
        proposed to incorporate into Bonsai has been developed by us, or we are
        experienced using it.

    \item[lack of interest from Bonsai users] We will minimize the probability
        of this problem by advertising broadly the new functionality upon
        release.  However, we don't think this would be a real problem, since
        Bonsai users have reapeatly asked the addition of machine learning
        functionality to Bonsai (Section 3 Community demand, Case for Support),
        and because the machine learning functionality we proposed to add to
        Bonsai are popular among experimental-control neurocientists.

    \item[lack of interest from machine-learning methods developers] To
        minimize the probability that this risk turns into a large problem we
        will take three actions. First, we will provide machine-learning
        methods developers and easy-to-use API to interface their methods into
        the Bonsai ecosystem. Second, we will show machine-learning methods
        developers the large user exposure that Bonsai provides to integrated
        ML functionality. Third, we will organize Bonsai workshops for methods
        developers showcasing machine learning functionality integrated into
        the Bonsai ecosystem and instructing them how to integrate new methods
        into Bonsai.

@@

\subsection{Metadata standards and data documentation}

% \textcolor{orange}{\textit{What metadata is produced about the data generated from the research? For example descriptions of data that enable research data to be used by others outside of your own team.  This may include documenting the methods used to generate the data, analytical and procedural information, capturing instrument metadata alongside data, documenting provenance of data and their coding, detailed descriptions for variables, records, etc.}}

@@

% documentation
The documentation, including learning materials and all relevant community
links are publicly accessible at \url{bonsai-rx.org}. A forum where Bonsai
users daily exchange questions and answers can be accessed at the public
repository \url{https://groups.google.com/g/bonsai-users}. Currently more than
700 users contribute to this forum (Figure~1f, \emph{Case for Support}).

Classes, methods and functions will be documented in the code, and automatic
documentation tools (e.g., Sphinx for Python code, or Sphinx for .Net for C\#)
will be used to automatically generate API documentation. API documentation
will be complemented with class and interaction diagrams, documenting the
structure and dynamics of the application objects.

@@

\textcolor{orange}{\textit{Plans and place for long-term storage, preservation,
and planned retention period for the research data. Formal preservation
standards, if any. Indicate which data may not be retained (if any). Also,
briefly describe what will be stored long-term, raw data, processed data?}}

For persistent and long term preservation we propose to use
Zenodo\footnote{\url{https://zenodo.org/}}.

@@

Software modules and documentation will be available on Github, with links from
the Bonsai project site (https://bonsai-rx.org) and available through the
Bonsai system's package management interface.

NEED: github org and status?

@@

@@

\subsection{Restrictions or delays to sharing, with planned actions to limit such restrictions} 
Restriction to data sharing may be due to subject confidentiality or to ensure you can gain IP protection. Applicants should ensure they have obtained necessary clearances from relevant collaborators with regards to the content of the proposal including the data sharing plan in line with the BBSRC Research Grants Guide. Where possible please indicate when data will be made available following a period of restriction.

All distributed software will be developed as part of the project, based on published and unrestricted algorithms.  Thus it is unencumbered by IP restrictions.  Will be available as above.

4. Data security (where relevant)
4.1	Formal information/data security standards
% Identify formal information standards with which your study is or will be compliant. An example is ISO 27001. If your organisation is ISO compliant, please state the registration number.

Not applicable.

4.2	Main risks to data security
% All personal data has an element of risk. Summarise the main risks to the confidentiality and security of information related to human participants or users, the level of risk and how these risks will be managed. Cover the main processes or facilities for storage and processing of personal data, data access, with controls put in place and any auditing of user compliance with consent and security conditions. It is not sufficient to write ‘not applicable’ under this heading.

This is a freely-available open source software package, which does not contain participants or personal data.   User access to the repository is anonymous and user personal information will not be collected.

>>>>

Software modules will comprise source files in various programming languages:
C\#, Python, MATLAB and R.  These will be accompanied by documentation
(Section~\ref{sec:documentation}).

% **NEED**: Approximate lines of code??? (order of magnitude is fine)

% **NEED**: testing files, makefiles?, other stuff,

@@

% code
The source code for the core Bonsai ecosystem and the majority of extension
packages are hosted, with MIT license (i.e., no copyright and licencing
restrictions), at \url{github.com/bonsai-rx}

@@

Team responsibilities: Dr.~Lopes (Bonsai related software development, software
releases, organization of Bonsai workshops and co-supervision of research
software engineer), Dr.~Rapela (machine learning software development, testing
maching learning functionality with behavioral and neural recordings and
co-supervision of research software engineer), Prof.~Sahani (supervision on
maching learning and neural data analysis), Prof.~Mrsic-Flogel (superivision on
neuroscience topics and communication with the steering committee).

We will use the Kanban software development model, which both Dr.~Lopes and
Dr.~Rapela are already using in a collaborative project at the SWC/GCNU.

Weekly meetings between Dr.~Lopes, Dr~Rapela and the to-be-hired research
software engineer will be an integral component of the software development
team. For instance, these meetings will discuss software engineering
topics, perform code reviews, or practice paired programming. They will help
every member of the software development team be informed about all aspect of
the software development project, which will ensure that knowledge is not lost
if a team member leaves the project. In addition, code review and
pair-programming sessions will guarantee that the generated code shares a
uniform style, which will help new contributors to the project (e.g., machine
learning methods developers) easily understand the code functionality.

We speculate the following risks of the project:

@@

    \item[problem in language integration] we proposed to use existing software
        libraries to enable the communication between Bonsai, written in C\#,
        and other machine learning algorithms, written in Python (Python.NET),
        R (R.Net) and Matlab (native Matlab .Net interface). We have not yet
        tested any of these existing libraries, and we may found that they
        don't function reliably enough for our purposes. If this were the case,
        we proposed to build our own interface protocol using the ZeroMQ
        messaging library.  This risk will only delay the progress of the
        proposed project, but will not hinder the deployment of its proposed
        outputs.

    \item[disfunctional machine learning functionality] in Section 2.3 of the
        Case for Support document (Figure 2, green box) we proposed to
        integrate existing machine learning methods into the Bonsai ecsosystem.
        It could be that some of these methods do not work as reliably as
        needed for our resource. In such a case we will use another existing
        methods with similar functionality.  This risk should not be a major
        problem for the proposed project, since most of the machine learning
        methods we proposed to incorporate into Bonsai have been
        developed by us, or we are substantial experienced using them.

    \item[lack of interest from Bonsai users] we will minimize the probability
        of this problem by demonstrating broadly among Bonsai users the unique
        capabilities of the new machine learning functionality.  We don't think
        this would be a serious problem, since Bonsai users have reapeatly
        asked the addition of machine learning functionality to Bonsai (Section
        3 Community demand, Case for Support), and because the functionality we
        proposed to add to Bonsai (Section~2.3, Case for Support) is of large
        interest among experimental-control neurocientists.

    \item[lack of interest from machine-learning methods developers] to reduce
        the probability of this problem we will take three actions. First, we
        will provide machine-learning methods developers and easy-to-use API to
        interface their methods into the Bonsai ecosystem. Second, we will
        demonstrate them the large user exposure that Bonsai offers to
        integrated maching learning functionality. Third, we will organize
        Bonsai workshops for methods developers showcasing machine learning
        methods already integrated into the Bonsai ecosystem and instructing
        them on how to integrate new methods into the Bonsai ecosystem.

@@

\subsection{Metadata standards and data documentation}
\label{sec:documentation}
% \textcolor{orange}{\textit{What metadata is produced about the data generated from the research? For example descriptions of data that enable research data to be used by others outside of your own team.  This may include documenting the methods used to generate the data, analytical and procedural information, capturing instrument metadata alongside data, documenting provenance of data and their coding, detailed descriptions for variables, records, etc.}}

@@

% documentation
The documentation, including learning materials and all relevant community
links are publicly accessible at \url{bonsai-rx.org}. A forum where Bonsai
users daily exchange questions and answers can be accessed at the public
repository \url{https://groups.google.com/g/bonsai-users}. Currently more than
700 users contribute to this forum (Figure~1f, \emph{Case for Support}).
Documentation will be created for the new modules produced in this project that
meets the existing standards, and will be integrated into the Bonsai
documentation framework.

Classes, methods and functions will be documented in the code, and automatic
documentation tools (e.g., Sphinx for Python code, or Sphinx for .Net for C\#)
will be used to automatically generate API documentation. This documentation
will be complemented with class and interaction diagrams, documenting the
structure and dynamics of the application objects.

@@

% \textcolor{orange}{\textit{Plans and place for long-term storage, preservation, and planned retention period for the research data. Formal preservation standards, if any. Indicate which data may not be retained (if any). Also, briefly describe what will be stored long-term, raw data, processed data?}}

For persistent and long term preservation we propose to use
Zenodo\footnote{\url{https://zenodo.org/}}, which is well integrated with
GitHub.

@@

Software modules will be available on Github and on Zenodo for long-term
preservation, with links from the Bonsai project site
(\url{https://bonsai-rx.org}) and available through the Bonsai system's package
management interface. Tutorial on how to use the added machine learning
functionality will be available at the Bonsai project site
(\url{https://bonsai-rx.org}) and automatically generated API documentation for
this functionality will be available through
ReadTheDocs\footnote{\url{https://readthedocs.org/}}.

@@

% Package management
% Bonsai website
% Community fora
% Workshops
% Paper?
 % A forum where Bonsai
% users daily exchange questions and answers can be accessed at
% \url{https://groups.google.com/g/bonsai-users}. Currently more than 700 users
% contribute to this forum (Figure~1f, \emph{Case for Support}).

Currently Bonsai is a well-known software. A strategy that its author has
used to gain this visiability is to build interoperability with a large number
of open-source source projects for data acquistion. With the addition of the
proposed machine learning functionality, Bonsai will interact with popular
machine learning algorithms, which will confer Bonsai further visibility.

@@

Several factors will contribute to the reusability of new functionality
incorporated into Bonsai. Code implementing this functionality will be
available in Github and Zenodo, under MIT license. Emphasis on standard and
uniform code style, achieved with the help of code review and paired
programming, will generate readable code, easier to read and extend by third
parties. Automated build methods, will simplify code building and encourage
modification by new developers. Test cases with large coverage will motivate
software developers to modify Bonsai, without the risk of introducing
undetected bugs into the software. Carefully designed software, with simple and
clear object interfaces, will simplify the resuse of the proposed software.

\subsection{Restrictions or delays to sharing, with planned actions to limit such restrictions} 
% Restriction to data sharing may be due to subject confidentiality or to ensure you can gain IP protection. Applicants should ensure they have obtained necessary clearances from relevant collaborators with regards to the content of the proposal including the data sharing plan in line with the BBSRC Research Grants Guide. Where possible please indicate when data will be made available following a period of restriction.

All distributed software will be developed as part of the project, based on published and unrestricted algorithms.  Thus it is unencumbered by IP restrictions and will be available throughout development as described above.

\section{Data security (where relevant)}

The software produced will be freely available and open sourced.  It will contain no personal data.  User access to the repository will be anonymous and user personal information will not be collected.  Thus security concerns do not apply.
